---
title: "hw2"
author: "Dylan Asoh"
date: "9/27/2021"
output: html_document
---
https://github.com/dylanasoh/stat433/tree/master/hw2

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cars}
library(nycflights13)
library(dplyr)
library(ggplot2)
head(flights)
```
```{r}
flights %>% filter(is.na(dep_time))
```

1. There are 8255 flights with a missing dep_time. The arrival time, departure delay, arrival delay, tail number, arrival delay, and air_time are also missing.These rows could represent cancelled flights.

2.

```{r}
flights <- flights %>% mutate(dep_time_since_midnight=(dep_time%/%100) * 60 + (dep_time%% 100),
                                                       sched_dep_time_since_midnight=(sched_dep_time%/%100) * 60 + (sched_dep_time%% 100))
flights %>% select(dep_time, dep_time_since_midnight, sched_dep_time, sched_dep_time_since_midnight)
```

```{r}
avg_delays_cancelled <- flights %>% 
  mutate(cancelled_flight=is.na(dep_time) | is.na(arr_time)) %>% 
  group_by(year, month, day) %>% 
  summarise(prop_cancelled=mean(cancelled_flight), total_flights=n(), avg_dep_delay=mean(dep_delay, na.rm=TRUE), avg_arr_delay=mean(arr_delay, na.rm=TRUE)) 
avg_delays_cancelled %>% ggplot(aes(avg_dep_delay,prop_cancelled))  + geom_point()
```

```{r}
avg_delays_cancelled %>% ggplot(aes(avg_arr_delay, prop_cancelled)) + geom_point()
```

3. There is a pattern between the number of cancelled flights per day and the average delay per day. There is a strong increasing relationship between the proportion of cancelled flights and average arrival delay. There is also a strong relationship between the proprotion of cancelled flights and the average departure delay as shown by the scatter plots.