---
title: "practice_midterm"
author: "Dylan Asoh"
output: html_document
---
  
## Midterm.
  
#### 1.  Map the delay by destination.
  
  Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States.  You are welcome to use this code or some other code. 

```{r} 
library(tidyverse)
library(nycflights13)
flights
```

You might want to use the size or colour of the points to display the average delay for each airport.

```{r}
new_flights <- flights %>% group_by(dest) %>% mutate(avg_delay=mean(arr_delay, na.rm=TRUE))
airports %>% inner_join(new_flights, c("faa" = "dest")) %>% ggplot(aes(lon, lat, color=avg_delay)) +
  borders("state") +
  geom_point() +
  coord_quickmap()
```


#### 2.   Do planes trade ownership?

You might expect that there’s an implicit relationship between plane and airline, because each plane is flown by a single airline. Explore this conjecture using data. (Let's assume that the tail number of a plane does not change.)

```{r}
# put your answer here. 
flights_airline <- flights %>% inner_join(airlines)
flights_airline %>% distinct(tailnum)
sum_distinct_planes_per_airline <- flights_airline %>% rename(airline=name) %>% group_by(airline) %>% summarise(distinct=n_distinct(tailnum))  
sum(sum_distinct_planes_per_airline$distinct)
```
The sum of the distinct planes per airline (4067) is greater than the number of distinct planes (4044). Therefore, 23 planes trade ownership.

#### 3a.  Plane's average speed.

Notice that ```flights$air_time``` is in minutes.  Make a new column that is the air time in hours. 

```{r}
flights_air_time_hours <- flights %>% mutate(air_time_hour=air_time / 60)
flights_air_time_hours
```


#### 4b. Average speed

For each flight, compute the average speed of that flight (in miles per hour).  Then, for each plane, compute the average of those average speeds. Display it in a histogram. You can use a base R histogram ```hist``` or ggplot's ```geom_histogram```. 

```{r}
flights_miles_per_hour <- flights_air_time_hours %>% mutate(mph=distance/air_time_hour)
flights_miles_per_hour %>% group_by(tailnum) %>% summarise(avg_mph=mean(mph, na.rm=TRUE)) %>% ggplot(aes(x=avg_mph))  + geom_histogram()
```




#### 5. Bonus

Make a table where each row is a destination, each column is a carrier, and each element is the number of times that the carrier has flown to that destination.  Ensure that you only count flights that arrived at the destination. 

```{r}
flights_airline %>% group_by(dest, carrier) %>% summarise(count=n()) %>% pivot_wider(names_from =carrier, values_from=count)
```
